---
title: "Seeing Democracy Through the Trees"
author: "Daniel K Baissa"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)

require(rJava)
# options(java.parameters = "-Xmx50g")     # or 8g, or larger than this, ...

options(java.parameters = "-Xmx8g")     # or 8g, or larger than this, ...
require(bartMachine)
set_bart_machine_num_cores(5)
```

Load in Vdem Data

```{r}
vdem <- readRDS("~/R Projects/Democracy_Prediction/Vdem_data/V-Dem-CY-Full+Others-v13.rds")
```

This dataset is the Country-Year: V-Dem Core



v2x_polyarchy will be our democracy indicator so the other higher level indicators will need to be removed.

v2x_egaldem
v2x_delibdem
v2x_partipdem
v2x_libdem

```{r}
vdem <- vdem |> 
  select(-c(v2x_egaldem, v2x_delibdem, v2x_partipdem, v2x_libdem))
```



Now let's look for clumns with NAs

```{r}
# Calculate the percentage of NAs in each column
na_percent <- colMeans(is.na(vdem)) * 100

# Find the columns with 20% or more NAs
columns_to_remove <- names(na_percent[na_percent >= 10])

# Remove the columns with 20% or more NAs from the dataset
vdem <- vdem[, !(names(vdem) %in% columns_to_remove)]

vdem <- vdem |> 
  select(year, v2x_partipdem_codelow, v2x_polyarchy)

```

Now let's break the data into a training set and a testing set. We will do this by randomly sampling 80% of the data into the training and 20% into the testing datasets.

```{r}
set.seed(123)
split_idx <- sample(1:nrow(vdem), floor(0.8 * nrow(vdem)))
train_data <- vdem[split_idx, ]
test_data <- vdem[-split_idx, ]

```


For fiting this BART model, I will do some basic data orginziation in order to make the process easily interpretible.

We will break the data into 2 components. 1) the y variable and 2) the x variables.

```{r}
train_data <- train_data |> 
  na.omit() 

y_train <- train_data$v2x_polyarchy

train_data$v2x_polyarchy <- NULL

test_data <- test_data |> 
  na.omit() 

y_test <- test_data$v2x_polyarchy

test_data$v2x_polyarchy <-NULL

```




Now let's fit the BART model

```{r}
bm <- bartMachine(y=y_train, X = train_data)

bm
```

Now Let's test out of sample.

```{r}
predictions <- predict(bm, test_data)
accuracy <- mean((predictions - y_test)^2)
cat("Test accuracy (MSE):", accuracy)
```

```{r}
vs <- var_selection_by_permute(bm,
                               num_permute_samples = 10,
                               bottom_margin = 1)
```

```{r}
vs <- var_selection_by_permute_cv(bm, k_folds = 2,
                                  num_permute_samples = 10)
```

```{r}
vdem_selected <- train_data |> 
  select(vs$important_vars_cv)

```


```{r}
bm_select <- bartMachine(y=y_train, X = train_data)

bm_select
```

```{r}
for (i in 1:length(vs$important_vars_cv)) {
  pd_plot(bm_select, j = vs$important_vars_cv[i])

}
```

